---
title: "p8105_hw2_jg4197"
author: "Jin Ge"
date: "9/25/2019"
output: github_document
---

## water health data tidiness

```{r problem 1, message = FALSE}
library(tidyverse)
library(readxl)

# help get relative paths 
getwd()

# read and clean Mr. Trash Wheel sheet
water_total_data <- 
  read_excel("E:./original_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
           sheet = "Mr. Trash Wheel",
           range = "A2:N338") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls),
         sports_balls = as.integer(sports_balls))
view(water_total_data)

# read and clean percipitation data for 2017 and 2018
water_17data <- read_excel("E:./original_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                           sheet = "2017 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  select(month, tl_precip = total) %>% 
  drop_na(month, tl_precip) %>% 
  mutate(year = "2017")

water_18data <- read_excel("E:./original_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                           sheet = "2018 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  select(month, tl_precip = total) %>% 
  drop_na(month, tl_precip) %>% 
  mutate(year = "2018")
view(water_17data)
view(water_18data)

# combine 2017 and 2018 precipitation datasets
comb_data <- left_join(water_17data, water_18data, by = "month") %>% 
  mutate(month = month.name)
view(comb_data)

# get water data in 2017
filt_17data <- filter(water_total_data, year == "2017")

view(filt_17data)


```
The total precipitation of 2018 is `r sum(pull(water_18data, tl_precip))`.

The total observations is `r count(water_total_data)`.

The number of 2017 and 2018 precipitation data is `r count(comb_data)`.

The key variables in the total dataset is `r names(water_total_data)`.

The median number of sports balls in 2017 is `r median(pull(filt_17data, sports_balls))`.


## politic and economic datasets tidiness

```{r problem 2}
# read and clean data in pols_month
pol_dat <- read_csv("E:./original_data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), 
           sep = c(4,7), convert = TRUE) %>%
  select(year, month, president = prez_dem) %>%  
  mutate(month = recode(month, `-01` = "January",
                        `-02` = "February",
                        `-03` = "March",
                        `-04` = "April",
                        `-05` = "May",
                        `-06` = "June",
                        `-07` = "July",
                        `-08` = "August",
                        `-09` = "September",
                        `-10` = "October",
                        `-11` = "November",
                        `-12` = "December"),
         president = recode(president, `0` = "gop",
                            `1` = "dem"))
view(pol_dat)


# read and clean data in snp
snp_dat <- read_csv("E:./original_data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>% 
  mutate(date = as.Date(date, "%m/%d/%Y")) %>% 
  separate(date, into = c("year", "month", "day"), sep = c(4,7)) %>% 
  select(year, month, snp_value = close) %>% 
  arrange(year, month) %>% 
  mutate(year = as.numeric(year),
    month = recode(month, `-01` = "January",
                        `-02` = "February",
                        `-03` = "March",
                        `-04` = "April",
                        `-05` = "May",
                        `-06` = "June",
                        `-07` = "July",
                        `-08` = "August",
                        `-09` = "September",
                        `-10` = "October",
                        `-11` = "November",
                        `-12` = "December")) 
view(snp_dat)

# read and clean unemployment data
unemply_dat <- read_csv("E:./original_data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemply_rate") %>% 
  mutate(month = recode(month, `jan` = "January",
                        `feb` = "February",
                        `mar` = "March",
                        `apr` = "April",
                        `may` = "May",
                        `jun` = "June",
                        `jul` = "July",
                        `aug` = "August",
                        `sep` = "September",
                        `oct` = "October",
                        `nov` = "November",
                        `dec` = "December"))
view(unemply_dat)

# join above three datasets
comb_3data <- left_join(pol_dat, snp_dat, by = c("year", "month")) %>% 
  left_join(unemply_dat, by = c("year", "month"))
view(comb_3data)
```

*pol_dat* contains the number of polls and give which party held the president position. *snp_dat* contains stock prices at closed points. *unemply_dat* contains unemployment rate each month of each year. 

The result dataset is a combined one, whose dimension is `r dim(comb_3data)`. 

The range of year in *comb_3data* is `r range(pull(comb_3data,year))`. 
The names of key variables are `r names(comb_3data)`.


## NYC popularity of baby names

```{r problem 3, message = FALSE}
# load and clean NYC baby name dataset
nyc_baby_name <- read_csv("E:./original_data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender),
         ethnicity = str_to_lower(ethnicity),
         ethnicity = replace(ethnicity, ethnicity == "asian and paci", "asian and pacific islander"),
         ethnicity = replace(ethnicity, ethnicity == "black non hisp", "black non hispanic"),
         ethnicity = replace(ethnicity, ethnicity == "white non hisp", "white non hispanic"),
         childs_first_name = str_to_title(childs_first_name)) %>% 
  distinct(childs_first_name, year_of_birth, gender, ethnicity, .keep_all = TRUE)
view(nyc_baby_name)

# name of olivia female babies table
olivia_female_tbl <- filter(nyc_baby_name, childs_first_name == "Olivia",
                     gender == "female") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(names_from = "ethnicity",
              values_from = "rank") 
view(olivia_female_tbl)

# popular male babies name table
popname_male_tbl <- filter(nyc_baby_name, rank == "1",
                        gender == "male") %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  pivot_wider(names_from = "ethnicity",
              values_from = "childs_first_name")
view(popname_male_tbl)


# scatterplot with count against rank in male white non-hispanic baby name
## create a table as a variable for plotting
male_whitnonhis_tbl <- filter(nyc_baby_name, ethnicity == "white non hispanic",
                              gender == "male", 
                              year_of_birth == "2016")
view(male_whitnonhis_tbl)

## coding for a scatterplot
malewhit_babname_plot <- ggplot(male_whitnonhis_tbl, aes(x = rank, y = count)) + geom_point(aes(color = childs_first_name), alpha = .5)

```
Since the number of scatter is too large to picture a plot

